\textsf{\textbf{Droplet Electro-Bouncing in Low-Gravity}}

### 2018-01-23 doc/manuscript/theory.tex
added complete, intro, theory, and methods sections

With the assumption of an intermediate range of Reynolds numbers $\mathbb{R}\mbox{e} \equiv \frac{2UR_d}{\nu}$, $1 \leq \mathbb{R}\mbox{e} \leq 5000 $ then the force drag drag acting on the droplet will be quadratic,
\begin{equation*}\label{drag_force}
\mathbf{F}_D = \frac{1}{2}C_D \rho A {y'}^2,
\end{equation*}
where $C_D$ is the drag coefficient, $\rho$ is the density of the sorrounding fluid medium (air, in this case), and $A$ is the frontal area of the droplet. For this range of Reynolds numbers it is appropriate to approximate the drag coefficient by the well known Abraham correlation \cite{abraham_functional_1970}
\[C_D = \frac{24}{9.06^2} \left( 1 + \frac{9.06}{\sqrt{\mathbb{R}\mbox{e}}} \right)^2 .\]

Modeling the electrostatic force is somehwhat more involved, but we will start with the standard electrohydrodynamic (EHD) approximation \cite{saville_electrohydrodynamics:taylor-melcher_1997} as our model, and further simplify where possible. Under a DC electric field, we assume that the real part of the dielectric permitivity $\epsilon$, $Re \langle \epsilon \rangle \approx  \mbox{constant}$. We also assume that electric currents are small enough that the effects of magnetic fields can be neglected. The validity of this assumption rests on the size of the characteristic time scale for electrical phenomena $\tau_e = \epsilon \epsilon_0/\sigma_e \ll 1$, where $\tau_e$ is the ratio of absolute dielectric permittivity $\kappa = \epsilon \epsilon_0$, to conductivity $\sigma_e$, of the medium. The time characteristic time $\tau_e$ is also known as the relaxation time, and is a measure of how quickly the polarization of a dielectric responds to a change in electric field. Given the respective conductivity, and permittivity of extremely-pure water ($ \kappa \sim 80$, $\sigma_e = 18.2 \times 10^{6}$ $\Omega\mbox{cm}$) \cite{yatsuzuka_electrification_1994}, we estimate $\tau_e \approx 4 \times 10^{-6}$ s. The relaxation for the common distilled water that is actually used in the droplet experiments is undobtedly even smaller (due to the presence of solvated ions). Neglecting the effects of a electric double layer on hydration of ions (in the water or the ambient atmosphere), then the assumption of small relaxation time further implies that the net charge present in the droplets will remain approximately constant during the typical time interval of a low-gravity experiment.

Supposing that electrical forces acting on free charges and dipoles in a fluid are transferred directly to the fluid itself, then this overall electrical body force will be the the divergence of the Maxwell stress tensor $\tau_m $, by

\[ \mathbf{F}_E = \nabla \cdot \tau_m = \nabla \cdot \left( \epsilon \epsilon_0 \mathbf{E} \mathbf{E} - \frac{1}{2} \epsilon \epsilon_0 \mathbf{E} \cdot \mathbf{E} \delta \right) ,\]
where $\mathbf{F}_E$ is the electric body force per unit volume, and $\delta$ is the delta function. The product of the electric field vectors is the dyadic product.

The classical Korteweg-Helmholtz force density formulation of the Maxwell stress tensor is usually expressed as \cite{melcher_continuum_????}
\begin{equation}\label{force_density}
\mathbf{F}_E = \rho_f \mathbf{E} + \frac{1}{2} \left| E \right|^2 \nabla \epsilon - \nabla \left( \frac{1}{2} \rho \left( \frac{\partial \epsilon}{\partial \rho} \right)_T \left| E \right|^2 \right) .
\end{equation}

The first term in this expression, equivalently written as $q\mathbf{E}$, is the well known Coulombic force or electrophoretic force, which arises from the presence of free charge in an external electric field. We expect this term to dominate the electric force in a DC field. The second term is the force arising from polarization stresses due to a nonuniform field acting across a gradient in permittivity. This force is widely termed the dielectrophoretic force (DEP). The third term describes forces due to electrostriction. It has been noted by Melcher and Hurwitz that the electrostriction term is the gradient of a scalar and can thus be cannononically lumped togeather with the hydrostatic pressure for incompressible fluids) \cite{hurwitz_electrohydrodynamic_1966}; we neglect it in our analysis.

It is common to approximate the polarization stress by idealizing the droplet as a simple dipole using the effective dipole moment method first suggested by Pohl in 1958 \cite{pohl_effects_1958}. This approach has been related back to the force density by the volumetric integration of the force density the substitution of a Taylor series expansion of $\mathbf{E}$ in the limit of a small radial gradient of the field within the dielectric sphere \cite{wang_general_1997}. The DEP force is related to the dipole moment (induced or permanent) of polarizable media which has a tendency to align the dipole with the electric field. If there is a gradient in the field then for a finite separation of charge one end of the dipole will feel a stronger electric field than the other, resulting in a net force. Whether the force is positive or negative in the direction of the electric field gradient depends on the difference of dielectric permittivites between the fluids, rather than on the polarity of $\mathbf{E}$ itself. In principle an external electric field will tend to induce a dipole in a dielectric material, but if the field is spatially uniform there is no gradient in the field the forces felt by the dipoles are symmetric and thus no net force. The dipole moment of a spherical linear-dielectric particle immersed in a dielectric medium is given by
\begin{equation} \label{dipole_m_1}
\mu = V_d \mathbf{P} = \frac{4}{3} \pi R_d^3 \mathbf{P},
\end{equation} 
where $\mathbf{P} = \left(\kappa_1 - 1 \right) \epsilon_0 \mathbf{E} = \chi_e \epsilon_0 \mathbf{E}$ is the polarization moment, and $R_d$ is the particle radius, $\kappa_1 = \frac{\epsilon}{\epsilon_0}$ being the relative dielectric constant of the external medium (air in this case), $\chi_e = \kappa_1 - 1$ being the electric susceptibility of the dielectric medium. The excess polarization $\mathbf{P}_e$, in the sphere is given by
\begin{equation} \label{polarization}
\mathbf{P}_e = \left( \kappa_2 - \kappa_1 \right) \epsilon_0 \mathbf{E} = \frac{3 \kappa_1}{\kappa_2 +2\kappa_1}\mathbf{E},
\end{equation}
where $\kappa_2$ is the relative dielectric constant of the spherical particle. Taking together Equations \ref{dipole_m_1}, and \ref{polarization} the effective dipole moment of the particle is given by 
\begin{equation}\label{dipole_m_2}
\mu = 4 \pi R_d^3 \left( \frac{\kappa_2 - \kappa_1}{\kappa_2 + 2 \kappa_1} \right) \kappa_1 \epsilon_0 \mathbf{E},
\end{equation}
and the force felt by the dipole is 
\begin{eqnarray} \label{dep_force}
\mathbf{F}_{DEP} &=& \left( \mathbf{P}_e \cdot \nabla \right) \mathbf{E} \nonumber \\
&=& 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2,
\end{eqnarray}
where it is nice to use the simplifying shorthand $K = \frac{\kappa_2 - \kappa_1}{\kappa_2 + 2 \kappa_1}$, which is also known as the Clausius-Mossotti factor. In cases where $K <$ 0, or $K>$ 0 the particle will be repelled or attracted to regions of strong field respectively. In our experiment, taking the relative dielectric constants to be $\kappa_1 \approx$ 1 and $\kappa_2 \approx$ 80, the Clausius-Mossotti factor $K \approx$ 0.96. It is important to note that the equivalent dipole approximation critically requires an assumption of small physical scale of the particle relative to the lengthscale of nonuniformity of the field, which in this case we take to be the length of the charged superhydrophobic surface ($L =$ 25 mm $\gg a \approx$ 2.5 mm).

When the droplet is close to the dielectric surface, the net charge on the droplet will tend to induce polarization of the dielectric, perturbing the electric field. The polarization bound charge in the dielectric will be of the opposite sign of the net droplet charge and thus there will be a force of attraction. This so-called image force is a correction to the Colomb force due to the external electric field only, and can be found by the method of images \cite{david_j._griffiths_introduction_1999}. The image force $\mathbf{F}_I$, is given by

\begin{equation}
\mathbf{F}_I = \frac{k q^2}{16 \pi \epsilon_0} y^{-2} \hat{\mathbf{j}},
\label{image_force}
\end{equation}
where the factor $k$ is a function of the dielectric surface susceptibility $k = \frac{\chi_e}{\chi_e + 2}$, and $\hat{\mathbf{j}}$ is a unit vector normal to the dielectric surface.

By substituting Equations \ref{dep_force}, \ref{image_force} into Equation \ref{force_density} we have
\begin{eqnarray*}
 \mathbf{F}_E &=& q \mathbf{E} + \mathbf{F}_{DEP} + \mathbf{F}_I \\
 &=& q \mathbf{E} + \frac{k q^2}{16 \pi \epsilon_0 } y^{-2} \hat{\mathbf{j}} + 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2, 
\end{eqnarray*}

and the 1-D governing equation becomes
\begin{eqnarray} \label{gov_eqn_subs}
&m y'' = - \frac{1}{2} C_D \rho A {y'}^2 - q E - \frac{k q^2}{16 \pi \epsilon_0} y^{-2}- 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2,& \nonumber \\
&y(0) = R, \hspace{1 mm} y'(0) = U_0 .&
\end{eqnarray}

By comparing DEP and Coulombic terms in Equation \ref{gov_eqn_subs}, we note that a condition to neglect the DEP term is
\begin{eqnarray}
1 &\gg& \frac{R_d^2 \kappa_2 \epsilon_0 K E_0}{q} \nonumber
\end{eqnarray}
As this condition likely holds in our experiments we henceforth neglect the DEP force in our analysis. There is some physical intuition to support this conclusion as well. The dielectric displacement, $\mathbf{D} = \kappa \epsilon_0 \mathbf{E}$ of a water droplet in air is very high due to its large relative dielectric constant. This implies that the field strength within the droplet is about 2 orders of magnitude smaller than in the surrounding medium (which is essentially the same as a vacuum from a dielectric standpoint), thus it is not particularly inaccurate to model the dielectric volume of a droplet as a conductive shell (an equipotential), with zero field in its interior.

\subsection{The Electric Field}
If we consider the charged dielectric surface of our experiments to be a square sheet of charge lying in the $xz$-plane with width $L$, the symmetry of the problem happily lets us obtain the $y$-component of the electric field $\mathbf{E}$ by direct integration. In particular it is easy to constuct the electric field due to a finite plane of charge by supoerposition of the electric fields of a series of line charges. By symmetry the electric field points along the $y$-axis; for a point along the $y$-axis the position vector is $\mathbf{r} = \left( x^2 + y^2 \right)^{1/2} \hat{\mathbf{r}}$ to the center of a line charge in the $xz$-plane. The $y$-component of $d\mathbf{E}$ is found by $d E_r = d E_y \cos \theta = d E_y y/ r$, where $\theta$ is the angle made between the $y$-axis and the positon vector $\mathbf{r}$. If the charge in a line element, $dx$ is $\sigma L dx$, where $\sigma$ is the surface charge density, the electric field of a line charge is given by \cite{david_j._griffiths_introduction_1999}
\[d E_r = \frac{k \sigma L dx}{4 \pi \epsilon_0 r \sqrt{r^2 + L^2/4}}.
\]
The $y$-component of the electric field $E_y$, is then
\[ E_y = \frac{\sigma L y }{4 \pi \epsilon_0} \int^{L/2}_{-L/2} \frac{1}{(y^2 + x^2) \sqrt{y^2 + x^2 + L^2/4}} dx 
.\]
With some substitutions this can be integrated to obtain an expression for the electric field in terms of $y$, 
\begin{equation}
E_y = \frac{\sigma}{ \pi \epsilon_0} \tan^{-1} \left( \frac{L^2}{y \sqrt{2L^2 + y^2}}\right)
.\end{equation}

By taking Taylor series expansions in large and small limits we can intuit a bit about the behavior of this field. In the limit $L \rightarrow \infty, \hspace{1mm} y \ll L$ the argument of the function tends towards infinity and
\[ \lim_{x\to\infty} \tan^{-1}(x) = \frac{\pi}{2}
,\]
and thus
\begin{equation}
E_y \approx \frac{\sigma}{4 \pi \epsilon_0} = E_0 \hspace{10mm} y \ll L
,\end{equation}
where $E_0$ is the characteristic electric field given by $E_0 = \frac{\sigma}{4 \pi \epsilon_0}$. This field is constant, and equivalent to the electric field due to an infinite plane of charge. In the limit of $y \gg L$, the argument of the arctangent function can be approximated by
\[ \frac{L^2}{2 y \left( 2 L^2 + 4 y^2\right)^{1/2}} = \frac{L^2}{4 y^2 \left( 1 + L^2/2y^2 \right)^{1/2}} \approx \frac{L^2}{4 y^2}
.\]
For small $x$, $\tan^{-1}(x) \sim x$ and we thus find the familiar electric field due to a point charge
\begin{equation}
E_y \approx L^2 E_0 y^{-2}  \hspace{10mm} y \gg L
.\end{equation}
Both these regimes can be clearly seen in the plot of $E_y$ shown in Figure \ref{fig:E0}.
\begin{figure}[htb]
    \centering
    \input{../figures/E0.pgf}
    \caption{A log-log plot of the magnitude of the non-dimensional electric field, $E_y$.\label{fig:E0}}
\end{figure}

\subsection{Scaling}
Introducing scaled variables
\begin{equation}
 \bar{t} = \frac{t}{t_c}, \hspace{10 mm} \bar{y} = \frac{y}{y_c}, \end{equation}
where $y_c$ and $t_c$ are characteristic length and time scales respectively, and using the cooridnate transformation $y(0) - R = 0$, the governing equation becomes
\begin{eqnarray}
& \bar{y}'' = - \frac{1}{2} \frac{C_D \rho A y_c}{m} \bar{y}'^2
- \frac{q E_0 t_c^2}{m y_c} \bar{E} 
- \frac{k q^2 t_c^2}{16 \pi \epsilon_0 R^2 m y_c} \left( \frac{y_c}{R} \bar{y} + 1 \right)^{-2}, & \nonumber \\
& \bar{y}(0) = 0, \hspace{1 mm} \bar{y}'(0) = \frac{U_0 t_c}{y_c} .&
\end{eqnarray}
We note several dimensionless groups
\[ \mathbf{\Pi}_1 = \frac{C_D \rho A y_c}{2 m}, \hspace{5 mm}
\mathbf{\Pi}_2 = \frac{q E_0 t_c^2}{m y_c}, \hspace{5 mm}
\mathbf{\Pi}_3 = \frac{k q^2 t_c^2}{16 \pi \epsilon_0 R^2 m y_c}, \hspace{5 mm}
\mathbf{\Pi}_4 = \frac{y_c}{R}.\]

\subsubsection{Inertial Electro-Image Limit}
In the limit of small $y, \hspace{1mm} t$ we expect intertia to scale as Coulombic and image force. With $y_c \sim U_0 t_c$ and picking $t_c$ such that Coulombic force is $\mathcal{O}(1)$
\[ t_c \sim \frac{m U_0}{q E_0}, \hspace{5 mm}
y_c \sim \frac{m U_0^2}{q E_0} .
\]
With these scales the governing equation then becomes
\begin{eqnarray}
& \bar{y}'' = -1 - \mathbb{I}\mbox{m} \left( \mathbb{E}\mbox{u}\bar{y} + 1 \right)^{-2} ,& \nonumber \\
& \bar{y}(0) = 0, \hspace{1 mm} \bar{y}'(0) = 1 .& \label{img_limit}
\end{eqnarray} 
with 
\[ \mathbb{I}\mbox{m} \equiv \frac{k q}{16 \pi \epsilon_0 R_d^2 E_0} = \mathbf{\Pi}_3, \hspace{5 mm}
\mathbb{E}\mbox{u} \equiv \frac{m U_0^2}{q E_0 R_d} = \mathbf{\Pi}_4 ,
\]
where $\mathbb{I}\mbox{m}$ is the Image number, and denotes the ratio of image forces to the Coulombic force of the unperturbed field, and where where $\mathbb{E}\mbox{u}$ is the electrostatic Euler number, and is a ratio of inertia to electrostatic force.

\subsubsection{Inertial Electro-Viscous Limit}
In the limit of large $y, \hspace{1mm} t$ we expect droplet inertia to scale as Coulombic force and drag. In this case there are several obvious choices of scales:
\begin{enumerate}
\item $y_c \sim U_0 t_c$ and make Coulomb force $\mathcal{O}(1)$.
\item $y_c \sim R_d$ or $L$, $t_c \sim \frac{L}{U_0}$ but this makes the governing equation singular.
\item $y_c \sim R_d$ or $L$, $t_c \sim \left( \frac{L m}{q E_0} \right)^{1/2}$.
\item $y_c \sim R_d$ or $L$, and making Coulomb force $\mathcal{O}(1)$.
\end{enumerate}

In Case 3, the characterisitc time is $t_c \sim \left( \frac{4 \pi R_d^2}{q E_0 L} \right)^2 $ and the non-dimensional governing equation is given by
\begin{eqnarray}
&\bar{y}'' = - \frac{C_D \rho A L}{2 m} \bar{y}'^2 - \left( \frac{L}{R_d} \bar{y} + 1\right)^{-2} ,& \nonumber \\
& \bar{y}(0) = \frac{R_d}{L}, \hspace{5 mm} \bar{y}'(0) = \left( \frac{4 \pi U_0^2 R_d^2}{q E_0 L^3}\right)^{1/2} = \frac{R_d}{L} \sqrt{\mathbb{E}\mbox{u}_+}.& \nonumber
\end{eqnarray}
where $\mathbb{E}\mbox{u}_+ = \frac{4 \pi m U_0^2}{q E_0 L}$ is a long time scaled electrostatic Euler number. We prefer the approach with the greatest physical simplicity, fewest Pi terms, and has homogoenous initial conditions.

>>>>

With the assumption of an intermediate range of Reynolds numbers $\mathbb{R}\mbox{e} \equiv \frac{2UR_d}{\nu}$, $1 \leq \mathbb{R}\mbox{e} \leq 5000 $ then the force of drag acting on the droplet will be quadratic,
\begin{equation*}\label{drag_force}
\mathbf{F}_D = \frac{1}{2}C_D \rho A {y'}^2,
\end{equation*}
where $C_D$ is the drag coefficient, $\rho$ is the density of the sorrounding fluid medium (air, in this case), and $A$ is the frontal area of the droplet. For this range of Reynolds numbers it is appropriate to approximate the drag coefficient by the well known Abraham correlation \cite{abraham_functional_1970}
\[C_D = \frac{24}{9.06^2} \left( 1 + \frac{9.06}{\sqrt{\mathbb{R}\mbox{e}}} \right)^2 .\]

Modeling the electrostatic force is somehwhat more involved, but we will start with the standard electrohydrodynamic (EHD) approximation \cite{saville_electrohydrodynamics:taylor-melcher_1997} as our model, and further simplify where possible. Under a DC electric field, we assume that the real part of the dielectric permitivity $\epsilon$, $Re \langle \epsilon \rangle \approx  \mbox{constant}$. We also assume that electric currents are small enough that the effects of magnetic fields can be neglected. The validity of this assumption rests on the size of the characteristic time scale for electrical phenomena $\tau_e = \epsilon \epsilon_0/\sigma_e \ll 1$, where $\tau_e$ is the ratio of absolute dielectric permittivity $\kappa = \epsilon \epsilon_0$, to conductivity $\sigma_e$, of the medium. This characteristic time $\tau_e$ is also known as the relaxation time, and is a measure of how quickly the polarization of a dielectric responds to a change in electric field. Given the respective conductivity, and permittivity of extremely-pure water ($ \kappa \sim 80$, $\sigma_e = 18.2 \times 10^{6}$ $\Omega\mbox{cm}$) \cite{yatsuzuka_electrification_1994}, we estimate $\tau_e \approx 4 \times 10^{-6}$ s. The relaxation time for the common distilled water that is actually used in the droplet experiments is undoubtedly even shorted (due to the presence of solvated ions). Neglecting the effects of a electric double layer on hydration of ions (in the water or the ambient atmosphere) due to the reletively massive size of the droplets studied, then the assumption of small relaxation time further implies that the free charge present in the droplets will remain approximately constant during the typical time interval of a low-gravity experiment.

Supposing that electrical forces acting on free charges and dipoles in a fluid are transferred directly to the fluid itself, then this overall electrical body force will be the the divergence of the Maxwell stress tensor $\tau_m $, by
\[ \mathbf{F}_E = \nabla \cdot \tau_m = \nabla \cdot \left( \epsilon \epsilon_0 \mathbf{E} \mathbf{E} - \frac{1}{2} \epsilon \epsilon_0 \mathbf{E} \cdot \mathbf{E} \delta \right) ,\]
where $\mathbf{F}_E$ is the electric body force per unit volume, and $\delta$ is the delta function. The product of the electric field vectors is the dyadic product.

The classical Korteweg-Helmholtz force density formulation of the Maxwell stress tensor is usually expressed as \cite{melcher_continuum_????}
\begin{equation}\label{force_density}
\mathbf{F}_E = \rho_f \mathbf{E} + \frac{1}{2} \left| E \right|^2 \nabla \epsilon - \nabla \left( \frac{1}{2} \rho \left( \frac{\partial \epsilon}{\partial \rho} \right)_T \left| E \right|^2 \right) .
\end{equation}
The first term in this expression, equivalently written as $q\mathbf{E}$, is the well known Coulombic force or electrophoretic force, which arises from the presence of free charge in an external electric field. We expect this term to dominate the electric force in a DC field. The second term is the force arising from polarization stresses due to a nonuniform field acting across a gradient in permittivity. This force is widely termed the dielectrophoretic force (DEP). The third term describes forces due to electrostriction. It has been noted by Melcher and Hurwitz that the electrostriction term is the gradient of a scalar and can thus be cannononically lumped togeather with the hydrostatic pressure for incompressible fluids) \cite{hurwitz_electrohydrodynamic_1966}; we neglect it in our analysis.

It is common to approximate the polarization stress by idealizing the droplet as a simple dipole using the effective dipole moment method first suggested by Pohl in 1958 \cite{pohl_effects_1958}. This approach has been related back to the force density by the volumetric integration of the force density with the substitution of a Taylor series expansion of $\mathbf{E}$ in the limit of a small radial gradient of the field within the dielectric sphere \cite{wang_general_1997}. The DEP force is related to the dipole moment (induced or permanent) of polarizable media which has a tendency to align the dipole with the electric field. If there is a gradient in the field then for a finite separation of charge one end of the dipole will feel a stronger electric field than the other, resulting in a net force. Whether the force is positive or negative in the direction of the electric field gradient depends on the difference of dielectric permittivites between the fluids, rather than on the polarity of $\mathbf{E}$ itself. In principle an external electric field will tend to induce a dipole in a dielectric material, but if the field is spatially uniform there is no gradient in the field and the forces felt by the dipoles are symmetric and thus there is no net force. The dipole moment of a spherical linear-dielectric particle immersed in a dielectric medium is given by
\begin{equation} \label{dipole_m_1}
\mu = V_d \mathbf{P} = \frac{4}{3} \pi R_d^3 \mathbf{P},
\end{equation} 
where $\mathbf{P} = \left(\kappa_1 - 1 \right) \epsilon_0 \mathbf{E} = \chi_e \epsilon_0 \mathbf{E}$ is the polarization moment, and $R_d$ is the particle radius, $\kappa_1 = \frac{\epsilon}{\epsilon_0}$ being the relative dielectric constant of the external medium (air in this case), $\chi_e = \kappa_1 - 1$ being the electric susceptibility of the dielectric medium. The excess polarization $\mathbf{P}_e$, in the sphere is given by
\begin{equation} \label{polarization}
\mathbf{P}_e = \left( \kappa_2 - \kappa_1 \right) \epsilon_0 \mathbf{E} = \frac{3 \kappa_1}{\kappa_2 +2\kappa_1}\mathbf{E},
\end{equation}
where $\kappa_2$ is the relative dielectric constant of the spherical particle. Taking together Equations \ref{dipole_m_1}, and \ref{polarization} the effective dipole moment of the particle is given by 
\begin{equation}\label{dipole_m_2}
\mu = 4 \pi R_d^3 \left( \frac{\kappa_2 - \kappa_1}{\kappa_2 + 2 \kappa_1} \right) \kappa_1 \epsilon_0 \mathbf{E},
\end{equation}
and the force felt by the dipole is 
\begin{eqnarray} \label{dep_force}
\mathbf{F}_{DEP} &=& \left( \mathbf{P}_e \cdot \nabla \right) \mathbf{E} \nonumber \\
&=& 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2,
\end{eqnarray}
where it is nice to use the simplifying shorthand $K = \frac{\kappa_2 - \kappa_1}{\kappa_2 + 2 \kappa_1}$, which is also known as the Clausius-Mossotti factor. In cases where $K <$ 0, or $K>$ 0 the particle will be repelled or attracted to regions of strong field respectively. In our experiment, taking the relative dielectric constants to be $\kappa_1 \approx$ 1 and $\kappa_2 \approx$ 80, the Clausius-Mossotti factor $K \approx$ 0.96. It is important to note that the equivalent dipole approximation critically requires an assumption of small physical scale of the particle relative to the lengthscale of nonuniformity of the field, which in this case we take to be the length of the charged superhydrophobic surface ($L =$ 25 mm $\gg a \approx$ 2.5 mm).

When the droplet is close to the dielectric surface, the free charge on the droplet will tend to induce polarization of the dielectric, perturbing the electric field. The polarization bound charge in the dielectric will be of the opposite sign of the free droplet charge and thus there will be a force of attraction. This so-called image force is a correction to the Colomb force due to the external electric field only, and can be found by the method of images \cite{david_j._griffiths_introduction_1999}. The image force $\mathbf{F}_I$, is given by
\begin{equation}
\mathbf{F}_I = \frac{k q^2}{16 \pi \epsilon_0} y^{-2} \hat{\mathbf{j}},
\label{image_force}
\end{equation}
where the factor $k$ is a function of the dielectric surface susceptibility $k = \frac{\chi_e}{\chi_e + 2}$, and $\hat{\mathbf{j}}$ is a unit vector normal to the dielectric surface.

By substituting Equations \ref{dep_force}, \ref{image_force} into Equation \ref{force_density} we have
\begin{eqnarray*}
 \mathbf{F}_E &=& q \mathbf{E} + \mathbf{F}_{DEP} + \mathbf{F}_I \\
 &=& q \mathbf{E} + \frac{k q^2}{16 \pi \epsilon_0 } y^{-2} \hat{\mathbf{j}} + 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2, 
\end{eqnarray*}
and the 1-D governing equation becomes
\begin{eqnarray} \label{gov_eqn_subs}
&m y'' = - \frac{1}{2} C_D \rho A {y'}^2 - q E - \frac{k q^2}{16 \pi \epsilon_0} y^{-2}- 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2,& \nonumber \\
&y(0) = R, \hspace{1 mm} y'(0) = U_0 .&
\end{eqnarray}

By comparing DEP and Coulombic terms in Equation \ref{gov_eqn_subs}, we note that a condition to neglect the DEP term is
\begin{eqnarray}
\frac{q}{ \kappa_2 \epsilon_0 K R_d^2 E_0} \gg 1 \nonumber
\end{eqnarray}
As this condition likely holds in our experiments we henceforth neglect the DEP force in our analysis. There is some physical intuition to support this conclusion as well. The dielectric displacement, $\mathbf{D} = \kappa \epsilon_0 \mathbf{E}$ of a water droplet in air is very high due to its large relative dielectric constant. This implies that the field strength within the droplet is about 2 orders of magnitude smaller than in the surrounding medium (which is essentially the same as a vacuum from a dielectric standpoint), thus it is not particularly inaccurate to model the dielectric volume of a droplet as a conductive shell (an equipotential), with zero field in its interior. As an aside, in treating the droplet as an ideal conductor we note that in our specific case the electrostatic force is not a body force \emph{per se} as the electric field is acting on charges on the surface of the conductor.

\subsection{The Electric Field}

If we consider the charged dielectric surface of our experiments to be a square sheet of charge lying in the $xz$-plane with width $L$, the symmetry of the problem happily lets us obtain the $y$-component of the electric field $\mathbf{E}$ by direct integration. In particular it is easy to constuct the electric field due to a finite plane of charge by supoerposition of the electric fields of a series of line charges. By symmetry the electric field points along the $y$-axis; for a point along the $y$-axis the position vector is $\mathbf{r} = \left( x^2 + y^2 \right)^{1/2} \hat{\mathbf{r}}$ to the center of a line charge in the $xz$-plane. The $y$-component of $d\mathbf{E}$ is found by $d E_r = d E_y \cos \theta = d E_y y/ r$, where $\theta$ is the angle made between the $y$-axis and the position vector $\mathbf{r}$. If the charge in a line element, $dx$ is $\sigma L dx$, where $\sigma$ is the average surface charge density, the electric field of a line charge is given by \cite{david_j._griffiths_introduction_1999}
\[d E_r = \frac{k \sigma L dx}{4 \pi \epsilon_0 r \sqrt{r^2 + L^2/4}}.
\]
The $y$-component of the electric field $E_y$, is then
\[ E_y = \frac{\sigma L y }{4 \pi \epsilon_0} \int^{L/2}_{-L/2} \frac{1}{(y^2 + x^2) \sqrt{y^2 + x^2 + L^2/4}} dx 
.\]
With some substitutions this can be integrated to obtain an expression for the electric field in terms of $y$, 
\begin{equation}
E_y = \frac{\sigma}{ \pi \epsilon_0} \tan^{-1} \left( \frac{L^2}{y \sqrt{2L^2 + y^2}}\right)
.\end{equation}
We note that this model of the electric field will be poor except in cases where $y \ll L, \hspace{1 mm} R_d \ll L$, which will be true for cases of large drops very close to the dielectric surface. For accurate models of the field in this case we likely would need to solve Laplace's equation for the electric potential $\varphi$
\[ - \nabla^2 \varphi = 0,\]
numerically using finite element methods.

By taking Taylor series expansions in large and small limits we can intuit a bit about the behavior of this field. In the limit $L \rightarrow \infty, \hspace{1mm} y \ll L$ the argument of the function tends towards infinity and
\[ \lim_{x\to\infty} \tan^{-1}(x) = \frac{\pi}{2}
,\]
and thus
\begin{equation}
E_y \approx \frac{\sigma}{4 \pi \epsilon_0} = E_0 \hspace{10mm} y \ll L
,\end{equation}
where $E_0$ is the characteristic electric field given by $E_0 = \frac{\sigma}{4 \pi \epsilon_0}$. This field is constant, and equivalent to the electric field due to an infinite plane of charge. In the limit of $y \gg L$, the argument of the arctangent function can be approximated by
\[ \frac{L^2}{2 y \left( 2 L^2 + 4 y^2\right)^{1/2}} = \frac{L^2}{4 y^2 \left( 1 + L^2/2y^2 \right)^{1/2}} \approx \frac{L^2}{4 y^2}
.\]
For small $x$, $\tan^{-1}(x) \sim x$ and we thus find the familiar electric field due to a point charge
\begin{equation}
E_y \approx L^2 E_0 y^{-2}  \hspace{10mm} y \gg L
.\end{equation}
Both of these regimes can be clearly seen in the plot of $E_y$ shown in Figure \ref{fig:E0}.
\begin{figure}[htb]
    \centering
    \input{../figures/E0.pgf}
    \caption{A log-log plot of the magnitude of the non-dimensional electric field, $E_y$.\label{fig:E0}}
\end{figure}

\subsection{Scaling}

The equation of motion is non-linear, non-homogenous and is best solved numerically, but nevertheless we are curious if we can obtain approximate solutions to the equation analytically using perturbation methods. This is aided by non-dimensionalizing the governing equations, and seeing if any dimensionless groups exist, and whether any are particualy small or large, and whether any fortuitous opportunities for simplifcation appear. Introducing the scaled variables
\begin{equation}
 \bar{t} = \frac{t}{t_c}, \hspace{10 mm} \bar{y} = \frac{y}{y_c}, \end{equation}
where $y_c$ and $t_c$ are characteristic length and time scales respectively, and using the cooridnate transformation $y(0) - R = 0$, the governing equation becomes
\begin{eqnarray}
& \bar{y}'' = - \frac{1}{2} \frac{C_D \rho A y_c}{m} \bar{y}'^2
- \frac{q E_0 t_c^2}{m y_c} \bar{E} 
- \frac{k q^2 t_c^2}{16 \pi \epsilon_0 R^2 m y_c} \left( \frac{y_c}{R} \bar{y} + 1 \right)^{-2}, & \nonumber \\
& \bar{y}(0) = 0, \hspace{1 mm} \bar{y}'(0) = \frac{U_0 t_c}{y_c} .& \label{scaled_eqn}
\end{eqnarray}
We note the existence of several dimensionless groups
\[ \mathbf{\Pi}_1 = \frac{C_D \rho A y_c}{2 m}, \hspace{5 mm}
\mathbf{\Pi}_2 = \frac{q E_0 t_c^2}{m y_c}, \hspace{5 mm}
\mathbf{\Pi}_3 = \frac{k q^2 t_c^2}{16 \pi \epsilon_0 R^2 m y_c}, \hspace{5 mm}
\mathbf{\Pi}_4 = \frac{y_c}{R}, \hspace{5 mm}
\mathbf{\Pi}_5 = \frac{U_0 t_c}{y_c}.\]

\subsubsection{Inertial Electro-Image Limit}

In the limit of small $y$ and $t$ we expect intertia to scale as Coulombic and image force. In this limit we can approximate the electric field as the constant, $E_0$. One possible characteristic lengthscale is $y_c \sim R_d$, however this ultimately is a poor choice. With $y_c \sim U_0 t_c$ and picking $t_c$ such that Coulombic force is $\mathcal{O}(1)$, the intrinsic scales become
\[ t_c \sim \frac{m U_0}{q E_0}, \hspace{5 mm}
y_c \sim \frac{m U_0^2}{q E_0} .
\]
With these scales the governing equation then becomes
\begin{eqnarray}
& \bar{y}'' = -1 - \mathbb{I}\mbox{m} \left( \mathbb{E}\mbox{u}\bar{y} + 1 \right)^{-2} ,& \nonumber \\
& \bar{y}(0) = 0, \hspace{1 mm} \bar{y}'(0) = 1 .& \label{img_limit}
\end{eqnarray} 
with 
\[ \mathbb{I}\mbox{m} \equiv \frac{k q}{16 \pi \epsilon_0 R_d^2 E_0} = \mathbf{\Pi}_3, \hspace{5 mm}
\mathbb{E}\mbox{u} \equiv \frac{m U_0^2}{q E_0 R_d} = \mathbf{\Pi}_4 ,
\]
where $\mathbb{I}\mbox{m}$ is the Image number, and denotes the ratio of image forces to the Coulombic force of the unperturbed field, and where $\mathbb{E}\mbox{u}$ is the electrostatic Euler number, and is a ratio of inertia to electrostatic force.

\subsubsection{Inertial Electro-Viscous Limit}
